Gammapy: A Python package for gamma-ray astronomy

### 2023-03-22 src/ms-review.tex

The \code{gammapy.irf} sub-package contains all classes and functionalities
to handle IRFs in a variety of formats.
Usually, \irfs store instrument properties in the form of multi-dimensional
tables, with quantities expressed in terms of energy (true or reconstructed),
off-axis angles or cartesian detector coordinates. The main quantities stored in
the common \gammaray \irfs are the effective area, energy dispersion,
PSF and background rate. The \code{gammapy.irf}
sub-package can open and access specific \irf extensions,
interpolate and evaluate the quantities of interest on both energy and spatial
axes, convert their format or units, plot or write them into
output files. In the following, we list the main classes of the
sub-package:

@@

\subsubsection{Energy Dispersion}
For \iacts, the energy resolution and bias, sometimes called energy dispersion,
is typically parametrised in terms of the so-called
migration parameter ($\mu$), which is defined as the ratio between the
reconstructed energy and the true energy. By definition, the mean of this ratio is
close to unity for a small energy bias and its distribution can 
be typically described by a Gaussian profile. However, more complex
shapes are also common. The migration parameter is given at each offset angle and
reconstructed energy. The main sub-classes are the \code{EnergyDispersion2D} which is
designed to handle the raw instrument description, and the \code{EDispKernelMap},
which contains an energy disperion matrix per sky position. I.e., a 4-dimensional
sky map where each position is associated to an energy dispersion matrix.
The energy dispersion matrix is a representation of the energy resolution
as a function of the true energy only and implemented in \gammapy
by the sub-class \code{EDispKernel}.

\subsubsection{Instrumental Background}
The instrumental background rate can be represented as either a 2-dimensional
data structure named \code{Background2D} or a 3-dimensional one named \code{Background3D}.
The background rate is stored as a differential count rate, normalised per solid angle
and energy interval at different reconstructed energies and offset angles.
In the \code{Background2D} case, the background is expected to follow a radially symmetric shape
and changes only with the offset angle from FoV center.
In the \code{Background3D} case, the background is allowed to vary with 
longitude and latitude of a tangential FoV coordinates system.

>>>>

The \code{gammapy.irf} sub-package contains all classes and functionalities
to handle IRFs in a variety of functional forms.
Usually, \irfs store instrument properties in the form of multi-dimensional
tables, with quantities expressed in terms of energy (true or reconstructed),
off-axis angles or cartesian detector coordinates. The main quantities stored in
the common \gammaray \irfs are the effective area, energy dispersion,
PSF and background rate. The \code{gammapy.irf}
sub-package can open and access specific \irf extensions,
interpolate and evaluate the quantities of interest on both energy and spatial
axes, convert their format or units, plot or write them into
output files. In the following, we list the main classes of the
sub-package:

@@

\subsubsection{Energy Dispersion}
For \iacts, the energy resolution and bias, sometimes called energy dispersion,
is typically parametrised in terms of the so-called
migration parameter ($\mu$, see section \label{sec:gammaray-data-analysis}).
 However, more complex
shapes are also common. The distribution of the migration parameter is given at each offset angle and
true energy. The main sub-classes are the \code{EnergyDispersion2D} which is
designed to handle the raw instrument description, and the \code{EDispKernelMap},
which contains an energy dispersion matrix per sky position, i.e., a 4-dimensional
sky map where each position is associated to an energy dispersion matrix.
The energy dispersion matrix is a representation of the energy resolution
as a function of the true energy only and implemented in \gammapy
by the sub-class \code{EDispKernel}.

\subsubsection{Instrumental Background}
The instrumental background rate can be represented as either a 2-dimensional
data structure named \code{Background2D} or a 3-dimensional one named \code{Background3D}.
The background rate is stored as a differential count rate, normalised per solid angle
and energy interval at different reconstructed energies and position in the FoV.
In the \code{Background2D} case, the background is expected to follow a radially symmetric shape
and changes only with the offset angle from FoV center.
In the \code{Background3D} case, the background is allowed to vary with 
longitude and latitude of a tangential FoV coordinates system.
